Weighted Model Counting with Conditional Weights

### 2020-08-04 paper/paper.tex
New template

\def\year{2020}\relax
%File: formatting-instruction.tex
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage{aaai20}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{graphicx}  % DO NOT CHANGE THIS
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
%\nocopyright
%PDF Info Is REQUIRED.
% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
 \pdfinfo{
/Title (AAAI Press Formatting Instructions for Authors Using LaTeX -- A Guide)
/Author (AAAI Press Staff, Pater Patel Schneider, Sunil Issar, J. Scott Penberthy, George Ferguson, Hans Guesgen)
} %Leave this	
% /Title ()
% Put your actual complete title (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case
% Leave the space between \Title and the beginning parenthesis alone
% /Author ()
% Put your actual complete list of authors (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case. 
% Each author should be only by a comma. If the name contains accents, remove them. If there are any LaTeX commands, 
% remove them.

% DISALLOWED PACKAGES
% \usepackage{authblk} -- This package is specifically forbidden
% \usepackage{balance} -- This package is specifically forbidden
% \usepackage{caption} -- This package is specifically forbidden
% \usepackage{color (if used in text)
% \usepackage{CJK} -- This package is specifically forbidden
% \usepackage{float} -- This package is specifically forbidden
% \usepackage{flushend} -- This package is specifically forbidden
% \usepackage{fontenc} -- This package is specifically forbidden
% \usepackage{fullpage} -- This package is specifically forbidden
% \usepackage{geometry} -- This package is specifically forbidden
% \usepackage{grffile} -- This package is specifically forbidden
% \usepackage{hyperref} -- This package is specifically forbidden
% \usepackage{navigator} -- This package is specifically forbidden
% (or any other package that embeds links such as navigator or hyperref)
% \indentfirst} -- This package is specifically forbidden
% \layout} -- This package is specifically forbidden
% \multicol} -- This package is specifically forbidden
% \nameref} -- This package is specifically forbidden
% \natbib} -- This package is specifically forbidden -- use the following workaround:
% \usepackage{savetrees} -- This package is specifically forbidden
% \usepackage{setspace} -- This package is specifically forbidden
% \usepackage{stfloats} -- This package is specifically forbidden
% \usepackage{tabu} -- This package is specifically forbidden
% \usepackage{titlesec} -- This package is specifically forbidden
% \usepackage{tocbibind} -- This package is specifically forbidden
% \usepackage{ulem} -- This package is specifically forbidden
% \usepackage{wrapfig} -- This package is specifically forbidden
% DISALLOWED COMMANDS
% \nocopyright -- Your paper will not be published if you use this command
% \addtolength -- This command may not be used
% \balance -- This command may not be used
% \baselinestretch -- Your paper will not be published if you use this command
% \clearpage -- No page breaks of any kind may be used for the final version of your paper
% \columnsep -- This command may not be used
% \newpage -- No page breaks of any kind may be used for the final version of your paper
% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
% \pagestyle -- This command may not be used
% \tiny -- This is not an acceptable font size.
% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference

@@

\usepackage{amsthm}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usepackage[backgroundcolor=lightgray]{todonotes}
\usepackage{complexity}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage[binary-units]{siunitx}
\usepackage[inline]{enumitem}
\usetikzlibrary{cd}
\usetikzlibrary{bayesnet}
\usetikzlibrary{calc}
\DeclareMathOperator{\im}{im}
\newtheorem{theorem}{Theorem}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

% The file aaai20.sty is the style file for AAAI Press 
% proceedings, working notes, and technical reports.
%
\setlength\titlebox{2.5in} % If your paper contains an overfull \vbox too high warning at the beginning of the document, use this
% command to correct it. You may not alter the value below 2.5 in
\title{Weighted Model Counting with Conditional Weights}
%Your title must be in mixed case, not sentence case. 
% That means all verbs (including short verbs like be, is, using,and go), 
% nouns, adverbs, adjectives should be capitalized, including both words in hyphenated terms, while
% articles, conjunctions, and prepositions are lower case unless they
% directly follow a colon or long dash
\author{Written by AAAI Press Staff\textsuperscript{\rm 1}\thanks{Primarily Mike Hamilton of the Live Oak Press, LLC, with help from the AAAI Publications Committee}\\ \Large \textbf{AAAI Style Contributions by
Pater Patel Schneider,} \\ \Large \textbf{Sunil Issar, J. Scott Penberthy, George Ferguson, Hans Guesgen}\\ % All authors must be in the same font size and format. Use \Large and \textbf to achieve this result when breaking a line
\textsuperscript{\rm 1}Association for the Advancement of Artificial Intelligence\\ %If you have multiple authors and multiple affiliations
% use superscripts in text and roman font to identify them. For example, Sunil Issar,\textsuperscript{\rm 2} J. Scott Penberthy\textsuperscript{\rm 3} George Ferguson,\textsuperscript{\rm 4} Hans Guesgen\textsuperscript{\rm 5}. Note that the comma should be placed BEFORE the superscript for optimum readability
2275 East Bayshore Road, Suite 160\\
Palo Alto, California 94303\\
publications20@aaai.org % email address must be in roman text type, not monospace or sans serif
}

\begin{document}
\maketitle

\begin{abstract}
AAAI creates proceedings, working notes, and technical reports directly from
electronic source furnished by the authors. To ensure that all papers in the
publication have a uniform appearance, authors must adhere to the following
instructions.
\end{abstract}

@@

\noindent Foo
\begin{itemize}
\item The main narrative
  \begin{enumerate}
  \item Putting weights on literals is restrictive.
  \item To overcome this, it is common practice to add more literals.
  \item This is indeed always possible but much slower.
  \item We show how we can define conditional weights on literals, allowing us
    to encode any measure.
  \item We demonstrate a specific example of this by presenting a new way to
    encode Bayesian networks into instances of WMC and adapting a WMC algorithm
    (ADDMC) to run on the new format.
  \item We show that this results in significantly faster inference.
  \item We show that our encoding results in asymptotically fewer literals and
    fewer ADDs.
  \end{enumerate}
\item Introduce
  \begin{itemize}
  \item WMC \cite{DBLP:conf/aaai/SangBK05}
  \item ADDMC \cite{DBLP:conf/aaai/DudekPV20} and ADDs
    \cite{DBLP:journals/fmsd/BaharFGHMPS97}
  \item A Bayesian network is a directed acyclic graph with random variables as
    vertices that defines a probability distribution over them. The full
    probability distribution can be expressed as the product of the conditional
    probability distributions of all random variables. For discrete Bayesian
    networks (and we only consider discrete networks in this paper), each such
    distribution can be represented by a table known as a conditional
    probability table (CPT).
  \end{itemize}
\item[F] What are the main claims, what are the main takeaways, intuitive [???]
  of theorems to follow.
\item Our work is, in spirit, similar to... using ROBDDs for Inference in
  Bayesian Networks with Troubleshooting as an Example
  \cite{DBLP:conf/uai/NielsenWJK00}: this is an interesting approach where
  deterministic parts of a BN are expressed as Boolean functions, i.e.,
  extracting the logical from the probabilistic. Maybe I can make a case that my
  encoding (the textual version) kind of does the same, even though the
  algorithm doesn't.
\item Replace things like `Darwiche (Darwiche, 2005)' with `Darwiche (2005)' if
  possible (without natbib).
\end{itemize}

@@

\paragraph{WMC for Bayesian networks.} Hitherto, four techniques have been
proposed for encoding Bayesian networks into instances of WMC. We will identify
them based on the initials of authors as well as publications years:
\texttt{d02} \cite{DBLP:conf/kr/Darwiche02}, \texttt{sbk05}
\cite{DBLP:conf/aaai/SangBK05}, \texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05},
and \texttt{cd06} \cite{DBLP:conf/sat/ChaviraD06}. Below we summarise the
observed performance differences among them. Sang et al.
\cite{DBLP:conf/aaai/SangBK05} claim that \texttt{sbk05} is a smaller encoding
than \texttt{d02} with respect to both the number of clauses and the number of
variables but provide no experimental comparison. Chavira and Darwiche
\cite{DBLP:conf/ijcai/ChaviraD05} compare \texttt{cd05} with \texttt{d02} by
measuring the time it takes to compile either encoding into an arithmetic
circuit (but do not measure inference time). The more recent encoding
\texttt{cd05} always compiles faster and results in a smaller arithmetic circuit
(as measured by the number of edges). In their subsequent paper, the same
authors perform two sets of experiments (that are relevant to this summary)
\cite{DBLP:conf/sat/ChaviraD06}. First, they compile \texttt{cd05} and
\texttt{cd06} encodings into d-DNNF (i.e., deterministic decomposable negation
normal form \cite{DBLP:journals/jancl/Darwiche01}), measuring both compilation
time and numbers of edges in the d-DNNF diagram. The results are mostly in
favour of \texttt{cd06}. Second, they compare the inference time of
\texttt{sbk05} run with Cachet \cite{DBLP:conf/sat/SangBBKP04} with the compile
times of \texttt{cd05} and \texttt{cd06}, but only on five (types of) instances.
In these experiments, \texttt{cd06} is always faster than \texttt{cd05}, while
the comparison with \texttt{sbk05} is mixed. Sometimes \texttt{cd06} is orders
of magnitude faster than \texttt{sbk05}, sometimes slightly slower. The
performance difference between \texttt{sbk05} and \texttt{cd05} is even harder
to judge: \texttt{sbk05} is better on three out of five instances and worse on
the remaining two. Based on this description, one would expect \texttt{cd06} to
be faster than both \texttt{cd05} and \texttt{sbk05}, both of which should be
faster than \texttt{d02}. The experiments in \cref{sec:experiments}, however,
strongly disagree with this prediction, showing that the quality of an encoding
depends strongly on the underlying search algorithm or compilation technique.

@@

\todo[inline]{Describe BA-PS duality without the table.}

% \begin{table}
%   \centering
%   \caption{A comparison of Boolean-algebraic (BA) and set-theoretic (ST)
%     concepts for $2^U$ for any set $U$}
%   \label{tbl:notation}
%   \begin{tabular}{lccl}
%     \toprule
%     BA name & BA symbol & ST symbol & ST name \\
%     \midrule
%     bottom & $\bot$ & $\emptyset$ & empty set \\
%     top & $\top$ & $U$ & \\
%     meet, and & $\land$ & $\cap$ & intersection \\
%     join, or & $\lor$ & $\cup$ & union \\
%     complement, not & $\neg$ & $^c$ & complement \\
%             & $\le$ & $\subseteq$ & subset relation, set inclusion \\
%     atom & & & singleton, unit set \\
%     \bottomrule
%   \end{tabular}
% \end{table}

@@

Let $\mathcal{L}$ be a propositional logic with atoms $a$ and $b$, and let $U =
\{ a, b \}$. Then $2^U$, the power set of $U$, is the set of all models of
$\mathcal{L}$, and $2^{2^U}$ is the set of all formulas. These sets can also be
represented as Boolean algebras---see \cref{tbl:notation} for an overview of
notational differences and \cref{tbl:notation_example} for examples of how
various elements can be represented in both notations. Most importantly, note
that the word \emph{atom} has completely different meanings in logic and in
Boolean algebras. An atom in $\mathcal{L}$ is an atomic formula, i.e., an
element of $U$, whereas an atom in a Boolean algebra is (in set-theoretic terms)
a singleton set. For instance, an atom in $2^{2^U}$ corresponds to a model of
$\mathcal{L}$, i.e., an element of $2^U$. Unless referring specifically to a
logic, we will use the algebraic definition of an atom while referring to
logical atoms as \emph{variables}. In the rest of the paper, for any set $U$, we
will use set-theoretic notation for $2^U$ and Boolean-algebraic notation for
$2^{2^U}$, except for (Boolean) atoms in $2^{2^U}$ that are denoted as $\{x\}$
for some model $x \in 2^U$.

@@

Let $U$ be a set. A \emph{measure} is a function $\mu\colon 2^{2^U} \to
\mathbb{R}_{\ge 0}$ such that $\mu(\bot) = 0$, and $\mu(a \lor b) = \mu(a) +
\mu(b)$ for all $a, b \in 2^{2^U}$ whenever $a \land b = \bot$. A \emph{weight
  function} is a function $\nu\colon 2^U \to \mathbb{R}_{\ge 0}$. A weight
function is \emph{factored} if $\nu = \prod_{x \in U} \nu_x$ for some functions
$\nu_x\colon 2^{\{x\}} \to \mathbb{R}_{\ge 0}$, $x \in U$. We say that a weight
function $\nu\colon 2^U \to \mathbb{R}_{\ge 0}$ \emph{induces} a measure
$\mu_\nu\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$ if
\begin{equation} \label{eq:induced_measure}
  \mu_\nu(x) = \sum_{\{u\} \le x} \nu(u).
\end{equation}
Finally, a measure $\mu\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$ is
\emph{factorable} if there exists a factored weight function $\nu\colon 2^U \to
\mathbb{R}_{\ge 0}$ that induces $\mu$.

\begin{lemma} \label{prop:measure}
  The function $\mu_\nu$, (from \cref{eq:induced_measure}), is a measure.
\end{lemma}
% \begin{proof}
%   Note that $\mu_\nu(\bot) = 0$ since there are no atoms below $\bot$. Let $a, b
%   \in 2^{2^{U}}$ be such that $a \land b = \bot$. By elementary properties of
%   Boolean algebras, all atoms below $a \lor b$ are either below $a$ or below
%   $b$. Moreover, none of them can be below both $a$ and $b$ because then they
%   would have to be below $a \land b = \bot$. Thus
%   \[
%     \mu_\nu(a \lor b) = \sum_{\{u\} \le a \lor b} \nu(u) = \sum_{\{u\} \le a}
%     \nu(u) + \sum_{\{u\} \le b} \nu(u) = \mu_\nu(a) + \mu_\nu(b)
%   \]
%   as required.
% \end{proof}

In this formulation, the process of calculating the value of $\mu_\nu(x)$ for
some $x \in 2^{2^U}$ with a given definition of $\nu$ is known as WMC.

@@

\todo[inline]{Introduce the subsection and give it some context.}

@@

\todo[inline]{Add the proof that one can always go from any measure to a
  factorable measure by adding more literals and say: Not all measures are
  factorable but it is common practice to add more literals to make the measure
  factorable. We show that it's always possible and yet it can be faster not to
  do it.}

@@

\subsection{Proof of Correctness}

@@

\begin{lemma} \label{lemma:cpt}
  Let $X \in \mathcal{V}$ be a random variable with parents $\mathrm{pa}(X) = \{ Y_1,
  \dots, Y_n \}$. Then $\mathrm{CPT}_X\colon 2^{\mathcal{E}^*(X)} \to
  \mathbb{R}_{\ge 0}$ is such that for any $x \in \im X$ and $(y_1, \dots, y_n)
  \in \prod_{i=1}^n \im Y_i$,
  \[
    \mathrm{CPT}_X (T) = \Pr(X = x \mid Y_1 = y_1, \dots, Y_n = y_n),
  \]
  where $T = \{ \lambda_{X=x} \} \cup \{ \lambda_{Y_i=y_i} \mid i = 1, \dots, n
  \}$.
\end{lemma}
% \begin{proof}
% If $X$ is binary, then $\mathrm{CPT}_X$ is a sum of $2\prod_{i=1}^n |\im
%   Y_i|$ terms, one for each possible assignment of values to variables $X, Y_1,
%   \dots, Y_n$. Exactly one of these terms is nonzero when applied to $T$, and
%   it is equal to $\Pr(X = x \mid Y_1 = y_1, \dots, Y_n = y_n)$ by definition.

%   If $X$ is not binary, then $\left( \sum_{i=1}^m [\lambda_{X = x_i}]
%   \right)(T) = 1$, and $\left( \prod_{i=1}^m \prod_{j=i+1}^m
%     (\overline{[\lambda_{X = x_i}]} + \overline{[\lambda_{X = x_j}]})
%   \right)(T) = 1$, so $\mathrm{CPT}_X(T) = \Pr(X = x \mid Y_1 = y_1,
%   \dots, Y_n = y_n)$ by a similar argument as before.
% \end{proof}

>>>>

\def\year{2021}\relax
\documentclass[letterpaper]{article} % DO NOT CHANGE THIS
\usepackage{aaai21}  % DO NOT CHANGE THIS
\usepackage{times}  % DO NOT CHANGE THIS
\usepackage{helvet} % DO NOT CHANGE THIS
\usepackage{courier}  % DO NOT CHANGE THIS
\usepackage[hyphens]{url}  % DO NOT CHANGE THIS
\usepackage{graphicx} % DO NOT CHANGE THIS
\urlstyle{rm} % DO NOT CHANGE THIS
\def\UrlFont{\rm}  % DO NOT CHANGE THIS
\usepackage{graphicx}  % DO NOT CHANGE THIS
\usepackage{natbib}  % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\usepackage{caption} % DO NOT CHANGE THIS AND DO NOT ADD ANY OPTIONS TO IT
\frenchspacing  % DO NOT CHANGE THIS
\setlength{\pdfpagewidth}{8.5in}  % DO NOT CHANGE THIS
\setlength{\pdfpageheight}{11in}  % DO NOT CHANGE THIS
%\nocopyright
%PDF Info Is REQUIRED.
% For /Author, add all authors within the parentheses, separated by commas. No accents or commands.
% For /Title, add Title in Mixed Case. No accents or commands. Retain the parentheses.
\pdfinfo{
/Title (Weighted Model Counting with Conditional Weights)
/Author ()
/TemplateVersion (2021.1)
} %Leave this
% /Title ()
% Put your actual complete title (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case
% Leave the space between \Title and the beginning parenthesis alone
% /Author ()
% Put your actual complete list of authors (no codes, scripts, shortcuts, or LaTeX commands) within the parentheses in mixed case.
% Each author should be only by a comma. If the name contains accents, remove them. If there are any LaTeX commands,
% remove them.

% DISALLOWED PACKAGES
% \usepackage{authblk} -- This package is specifically forbidden
% \usepackage{balance} -- This package is specifically forbidden
% \usepackage{color (if used in text)
% \usepackage{CJK} -- This package is specifically forbidden
% \usepackage{float} -- This package is specifically forbidden
% \usepackage{flushend} -- This package is specifically forbidden
% \usepackage{fontenc} -- This package is specifically forbidden
% \usepackage{fullpage} -- This package is specifically forbidden
% \usepackage{geometry} -- This package is specifically forbidden
% \usepackage{grffile} -- This package is specifically forbidden
% \usepackage{hyperref} -- This package is specifically forbidden
% \usepackage{navigator} -- This package is specifically forbidden
% (or any other package that embeds links such as navigator or hyperref)
% \indentfirst} -- This package is specifically forbidden
% \layout} -- This package is specifically forbidden
% \multicol} -- This package is specifically forbidden
% \nameref} -- This package is specifically forbidden
% \usepackage{savetrees} -- This package is specifically forbidden
% \usepackage{setspace} -- This package is specifically forbidden
% \usepackage{stfloats} -- This package is specifically forbidden
% \usepackage{tabu} -- This package is specifically forbidden
% \usepackage{titlesec} -- This package is specifically forbidden
% \usepackage{tocbibind} -- This package is specifically forbidden
% \usepackage{ulem} -- This package is specifically forbidden
% \usepackage{wrapfig} -- This package is specifically forbidden
% DISALLOWED COMMANDS
% \nocopyright -- Your paper will not be published if you use this command
% \addtolength -- This command may not be used
% \balance -- This command may not be used
% \baselinestretch -- Your paper will not be published if you use this command
% \clearpage -- No page breaks of any kind may be used for the final version of your paper
% \columnsep -- This command may not be used
% \newpage -- No page breaks of any kind may be used for the final version of your paper
% \pagebreak -- No page breaks of any kind may be used for the final version of your paperr
% \pagestyle -- This command may not be used
% \tiny -- This is not an acceptable font size.
% \vspace{- -- No negative value may be used in proximity of a caption, figure, table, section, subsection, subsubsection, or reference
% \vskip{- -- No negative value may be used to alter spacing above or below a caption, figure, table, section, subsection, subsubsection, or reference

@@

% The file aaai21.sty is the style file for AAAI Press
% proceedings, working notes, and technical reports.
%

\usepackage{amsthm}
\usepackage{thmtools,thm-restate}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage[ruled,vlined]{algorithm2e}
\usepackage[capitalise]{cleveref}
\usepackage{bm}
\usepackage{booktabs}
\usepackage[table]{xcolor}
\usepackage{tikz}
\usepackage{complexity}
\usepackage{subcaption}
\usepackage{multirow}
\usepackage[binary-units]{siunitx}
\usepackage[inline]{enumitem}
\usetikzlibrary{cd}
\usetikzlibrary{bayesnet}
\usetikzlibrary{calc}
\DeclareMathOperator{\im}{im}
\newtheorem{theorem}{Theorem}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}
\newtheorem{proposition}{Proposition}
\newtheorem{corollary}{Corollary}
\theoremstyle{definition}
\newtheorem{definition}{Definition}
\newtheorem{example}{Example}
\theoremstyle{remark}
\newtheorem*{remark}{Remark}

\title{Weighted Model Counting with Conditional Weights}
%\author{
%  Paulius Dilkas, \textsuperscript{\rm 1}
%  Vaishak Belle \textsuperscript{\rm 1,2} \\
%}
%\affiliations{
%    \textsuperscript{\rm 1} University of Edinburgh \\
%    \textsuperscript{\rm 2} Alan Turing Institute \\
%    p.dilkas@sms.ed.ac.uk, vaishak@ed.ac.uk
%}
\begin{document}

\maketitle

\begin{abstract} % TODO
  AAAI creates proceedings, working notes, and technical reports directly from
  electronic source furnished by the authors. To ensure that all papers in the
  publication have a uniform appearance, authors must adhere to the following
  instructions.
\end{abstract}

@@

Introduce WMC \cite{DBLP:conf/aaai/SangBK05}

Over the last fifteen years, WMC has been extended and generalised in many ways,
e.g., to handle continuous probability distributions
\cite{DBLP:conf/ijcai/BellePB15}, first-order probabilistic theories
\cite{DBLP:conf/ijcai/BroeckTMDR11,DBLP:journals/cacm/GogateD16}, and infinite
domains \cite{DBLP:conf/aaai/Belle17}. Furthermore, WMC has been generalised
from weights in the interval $[0, 1]$ or positive real numbers to an arbitrary
semiring, making it applicable to domains such as parsing, dynamic programming,
constraint programming, databases, networks, etc.
\cite{DBLP:journals/japll/KimmigBR17}. It has been applied to Bayesian networks
\cite{DBLP:conf/aaai/SangBK05,DBLP:books/daglib/0024906} and probabilistic
graphical models more generally \cite{DBLP:conf/ecsqaru/ChoiKD13}, probabilistic
logic programming languages \cite{DBLP:journals/tplp/FierensBRSGTJR15} and
probabilistic programs with discrete probability distributions
\cite{DBLP:journals/corr/abs-2005-09089}. Many WMC algorithms have been proposed
based primarily either on knowledge compilation
\cite{DBLP:conf/ecai/Darwiche04,DBLP:conf/ijcai/OztokD15,DBLP:conf/ijcai/LagniezM17}
or search \cite{DBLP:conf/aaai/SangBK05}.

The most well-studied version of WMC assigns weights to models based on
weights on literals, i.e., the weight of a model is the product of the weights
of all literals in it. This severely restricts what probability distributions
can be represented, and this limitation is usually overcome by increasing the
size of the problem, i.e., adding more literals.

We show that while this workaround is always possible, it can be significantly
more efficient to consider \emph{conditional weights} on literals. Similarly to
conditional probabilities, conditional weights allow us to represent
probabilities in a manner that is both efficient and interpretable. Not having
additional variables that are only needed to manipulate the representation into
a specific format helps with both efficiency and interpretability. Similarly to
Bayesian networks, the size of the representation scales with independence and
literal-weight WMC is a special case when everything is independent.

We demonstrate a specific example of this by presenting a new way to
encode Bayesian networks into instances of WMC and adapting a WMC algorithm
ADDMC to run on the new format. ADDMC \cite{DBLP:conf/aaai/DudekPV20} is a
recently-proposed algorithm for WMC based on manipulating functions on Boolean
algebras using an efficient representation for such functions known as
\emph{algebraic decision diagrams} (ADDs)
\cite{DBLP:journals/fmsd/BaharFGHMPS97}. We show that with a few minor
modifications the algorithm can be adapted to work on our new WMC format

Showing that the new encoding produces an asymptotically smaller problem (w.r.t.
the numbers of literals and ADDs) and results in faster inferences.

% TODO: an overview of all results, with forward references

@@

% TODO: introduce this section and explain why there's so much about BNs
A Bayesian network is a directed acyclic graph with random variables as vertices
that defines a probability distribution over them. The full probability
distribution can be expressed as the product of the conditional probability
distributions of all random variables. For discrete Bayesian networks (and we
only consider discrete networks in this paper), each such distribution can be
represented by a table known as a conditional probability table (CPT).

\paragraph{WMC for Bayesian networks.} Hitherto, four techniques have been
proposed for encoding Bayesian networks into instances of WMC. We will identify
them based on the initials of authors as well as publications years:
\texttt{d02} \cite{DBLP:conf/kr/Darwiche02}, \texttt{sbk05}
\cite{DBLP:conf/aaai/SangBK05}, \texttt{cd05} \cite{DBLP:conf/ijcai/ChaviraD05},
and \texttt{cd06} \cite{DBLP:conf/sat/ChaviraD06}. Below we summarise the
observed performance differences among them. \citet{DBLP:conf/aaai/SangBK05}
claim that \texttt{sbk05} is a smaller encoding than \texttt{d02} with respect
to both the number of clauses and the number of variables but provide no
experimental comparison. \citet{DBLP:conf/ijcai/ChaviraD05} compare
\texttt{cd05} with \texttt{d02} by measuring the time it takes to compile either
encoding into an arithmetic circuit (but do not measure inference time). The
more recent encoding \texttt{cd05} always compiles faster and results in a
smaller arithmetic circuit (as measured by the number of edges). In their
subsequent paper, the same authors perform two sets of experiments (that are
relevant to this summary) \cite{DBLP:conf/sat/ChaviraD06}. First, they compile
\texttt{cd05} and \texttt{cd06} encodings into d-DNNF (i.e., deterministic
decomposable negation normal form \cite{DBLP:journals/jancl/Darwiche01}),
measuring both compilation time and numbers of edges in the d-DNNF diagram. The
results are mostly in favour of \texttt{cd06}. Second, they compare the
inference time of \texttt{sbk05} run with Cachet \cite{DBLP:conf/sat/SangBBKP04}
with the compile times of \texttt{cd05} and \texttt{cd06}, but only on five
(types of) instances. In these experiments, \texttt{cd06} is always faster than
\texttt{cd05}, while the comparison with \texttt{sbk05} is mixed. Sometimes
\texttt{cd06} is orders of magnitude faster than \texttt{sbk05}, sometimes
slightly slower. The performance difference between \texttt{sbk05} and
\texttt{cd05} is even harder to judge: \texttt{sbk05} is better on three out of
five instances and worse on the remaining two. Based on this description, one
would expect \texttt{cd06} to be faster than both \texttt{cd05} and
\texttt{sbk05}, both of which should be faster than \texttt{d02}. The
experiments in \cref{sec:experiments}, however, strongly disagree with this
prediction, showing that the quality of an encoding depends strongly on the
underlying search algorithm or compilation technique.

@@

@@

Let $\mathcal{L}$ be a propositional logic with atoms $a$ and $b$, and let $U =
\{ a, b \}$. Then $2^U$, the power set of $U$, is the set of all models of
$\mathcal{L}$, and $2^{2^U}$ is the set of all formulas. These sets can also be
represented as Boolean algebras (e.g., $(2^{2^U}, \land, \lor, \neg, \bot,
\top)$)---see \cref{tbl:notation_example} for examples of how various elements
can be represented in both notations. Most importantly, note that the word
\emph{atom} has completely different meanings in logic and in Boolean algebras.
An atom in $\mathcal{L}$ is an atomic formula, i.e., an element of $U$, whereas
an atom in a Boolean algebra is (in set-theoretic terms) a singleton set. For
instance, an atom in $2^{2^U}$ corresponds to a model of $\mathcal{L}$, i.e., an
element of $2^U$. Unless referring specifically to a logic, we will use the
algebraic definition of an atom while referring to logical atoms as
\emph{variables}. In the rest of the paper, for any set $U$, we will use
set-theoretic notation for $2^U$ and Boolean-algebraic notation for $2^{2^U}$,
except for (Boolean) atoms in $2^{2^U}$ that are denoted as $\{x\}$ for some
model $x \in 2^U$.

@@

Let $U$ be a set. A \emph{measure} is a function $\mu\colon 2^{2^U} \to
\mathbb{R}_{\ge 0}$ such that $\mu(\bot) = 0$, and $\mu(a \lor b) = \mu(a) +
\mu(b)$ for all $a, b \in 2^{2^U}$ whenever $a \land b = \bot$. A \emph{weight
  function} is a function $\nu\colon 2^U \to \mathbb{R}_{\ge 0}$. A weight
function is \emph{factored} if $\nu = \prod_{x \in U} \nu_x$ for some functions
$\nu_x\colon 2^{\{x\}} \to \mathbb{R}_{\ge 0}$, $x \in U$. We say that a weight
function $\nu\colon 2^U \to \mathbb{R}_{\ge 0}$ \emph{induces} a measure
$\mu_\nu\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$ if $\mu_\nu(x) = \sum_{\{u\} \le
  x} \nu(u)$.

\begin{restatable}{theorem}{measure} \label{prop:measure}
  The function $\mu_\nu$ is a measure.
\end{restatable}

Finally, a measure $\mu\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$ is
\emph{factorable} if there exists a factored weight function $\nu\colon 2^U \to
\mathbb{R}_{\ge 0}$ that induces $\mu$. In this formulation, the process of
calculating the value of $\mu_\nu(x)$ for some $x \in 2^{2^U}$ with a given
definition of $\nu$ is known as WMC.

@@

WMC with weights defined on literals is only able to capture a subset of all
possible measures on the Boolean algebra. This can be illustrated with a simple
example.

@@

Since many measures of interests may not be factorable, a well-known way to
encode them into instances of WMC is by adding more literals
\cite{DBLP:journals/ai/ChaviraD08}. We can indeed show that this is always
possible, however, as ensuing sections will demonstrate, it often makes the
inference task much harder.

\begin{restatable}{theorem}{extendable}
  For any set $U$ and measure $\mu\colon 2^{2^U} \to \mathbb{R}_{\ge 0}$, there
  exists a set $V \supseteq U$ and a factorable measure $\mu'\colon 2^{2^V} \to
  \mathbb{R}_{\ge 0}$ such that $\mu = \mu' \circ \iota$, where $\iota\colon
  2^{2^U} \to 2^{2^V}$ is the inclusion homomorphism.
\end{restatable}

@@

\subsection{Correctness}

@@

\begin{restatable}{lemma}{cptlemma} \label{lemma:cpt}
  Let $X \in \mathcal{V}$ be a random variable with parents $\mathrm{pa}(X) = \{ Y_1,
  \dots, Y_n \}$. Then $\mathrm{CPT}_X\colon 2^{\mathcal{E}^*(X)} \to
  \mathbb{R}_{\ge 0}$ is such that for any $x \in \im X$ and $(y_1, \dots, y_n)
  \in \prod_{i=1}^n \im Y_i$,
  \[
    \mathrm{CPT}_X (T) = \Pr(X = x \mid Y_1 = y_1, \dots, Y_n = y_n),
  \]
  where $T = \{ \lambda_{X=x} \} \cup \{ \lambda_{Y_i=y_i} \mid i = 1, \dots, n
  \}$.
\end{restatable}

\begin{restatable}{lemma}{fulldistribution} \label{lemma:full_distribution}
  Let $\mathcal{V} = \{X_1, \dots, X_n\}$. Then
  \[
    \phi(T) =
    \begin{cases}
      \Pr(x_1, \dots, x_n) &
      \begin{aligned}
        &\text{if } T = \{ \lambda_{X_i=x_i} \}_{i = 1}^n \text{ for} \\
        &\text{some } \textstyle (x_i)_{i=1}^n \in \prod_{i=1}^n \im X_i
      \end{aligned} \\
      0 & \text{otherwise,}
    \end{cases}
  \]
  for all $T \in 2^U$.
\end{restatable}

\begin{restatable}{theorem}{correctness} \label{thm:correctness}
  For any $X \in \mathcal{V}$ and $x \in \im X$,
  \[
    (\exists_U(\phi \cdot [\lambda_{X=x}]))(\emptyset) = \Pr(X = x).
  \]
\end{restatable}
